---
title: "ch9"
author: "mht"
date: "November 5, 2014"
output: html_document
---

# Chapter 9: Comparing binomial rates

## 9.1 Equality of proportions

```{r 9.1, echo=FALSE}

library(R2jags)
library(gridExtra)
library(reshape2)
library(polspline)
library(ggplot2)

setwd("/Users/mht/Documents/learning/tfbt/Lee&Wagenmakers/Code/ModelSelection/Means/")

cat('# Pledgers
model{
  # Rates and Difference
  theta1 ~ dbeta(1,1)
  theta2 ~ dbeta(1,1)
  delta <- theta1-theta2  
  # Data
  s1 ~ dbin(theta1,n1)
  s2 ~ dbin(theta2,n2)
  # Prior Sampling
  theta1prior ~ dbeta(1,1)
  theta2prior ~ dbeta(1,1)
  deltaprior <- theta1prior-theta2prior
}',file={f<-tempfile()})

s1 <- 424
s2 <- 5416
n1 <- 777
n2 <- 9072

# two-sided p-value = 0.005848:
prop.test(c(s1,s2), c(n1,n2), alternative = c("two.sided")) #approximate

# Analytical Bayes factor:
log.BF01 <- lchoose(n1,s1) + lchoose(n2,s2) + log(n1+1) + log(n2+1) - lchoose((n1+n2),(s1+s2)) - log(n1+n2+1)
BF01 <- exp(log.BF01)

data  <- list("s1","s2","n1","n2") # to be passed on to JAGS

myinits <- list(
  list(theta1 = runif(1), theta2 = runif(1), theta1prior = runif(1), theta2prior = runif(1)),
  list(theta1 = runif(1), theta2 = runif(1), theta1prior = runif(1), theta2prior = runif(1)),
  list(theta1 = runif(1), theta2 = runif(1), theta1prior = runif(1), theta2prior = runif(1)))

parameters <- c("theta1", "theta2", "delta", "deltaprior")

# The following command calls JAGS with specific options.
# For a detailed description see the R2jags documentation.
samples <- jags(data, inits=myinits, parameters,
   			model.file =f,
	 			n.chains=3, n.iter=10000, n.burnin=1000, n.thin=1, DIC=T)
# Now the values for delta and deltaprior are in the "samples" object, ready for inspection.

######################################################
# H1: delta is unrestricted
######################################################

# Collect posterior samples across all chains:
df <-data.frame(delta.post=samples$BUGSoutput$sims.list$delta,
                delta.prior=samples$BUGSoutput$sims.list$deltaprior)

#============ BFs based on logspline fit ===========================
library(polspline) # this package can be installed from within R
fit.prior     <- logspline(df$delta.prior, lbound=-1, ubound=1) # note the bounds.
fit.posterior <- logspline(df$delta.post, lbound=-1, ubound=1)

# 95% confidence interval:
x0 <- qlogspline(0.025,fit.posterior)
x1 <- qlogspline(0.975,fit.posterior)

posterior     <- dlogspline(0, fit.posterior) # this gives the pdf at point delta = 0
prior         <- dlogspline(0, fit.prior)     # based on the logspline fit
BF01          <- posterior/prior
# 1/BF01 gives 2.14 -- Exact solution: 2.223484
#BF01          <- posterior # because we know the height of the prior equals 1 at delta = 0 
# 1/BF01 gives 2.17```

a<-ggplot(data=df)+
  geom_density(aes(x=delta.post),linetype='dashed')+
  geom_density(aes(x=delta.prior))+
  geom_point(x=0,y=prior)+
  geom_point(x=0,y=posterior,color='red')+
  theme_bw()+
  xlim(-1,1)+
  xlab('Delta')

b<-ggplot(data=df)+
  geom_density(aes(x=delta.post),linetype='dashed')+
  geom_density(aes(x=delta.prior))+
  geom_point(x=0,y=prior)+
  geom_point(x=0,y=posterior,color='red')+
  theme_bw()+
  xlim(-0.05,0.05)+
  xlab('Delta')

grid.arrange(a,b,nrow=1)
```


Bayes factor is `r BF01` or 1 / `r BF01`

A reasonable interpretation of this Bayes factor is that the data do not pro- vide much evidence in favor of one hypothesis over the other. This seems more conservative than the interpretation that may be drawn from Bayesian parameter estimation, since the Bayesian 95% credible interval for the posterior of δ is approx- imately (−0.09,−0.01) and does not include 0. The reason for the discrepancy is that the Bayesian hypothesis test punishes H1 for assigning prior mass to values of δ that yield very low likelihoods.


### Exercise 9.1.1

Because the rate parameters θ1 and θ2 both have a uniform prior distribution, the prior distribution for the difference parameter δ can be found analytically as a triangular distribution. What are the advantages of using this result, rather than relying on computational sampling? What are the disadvantages?

Ans:

### Exercise 9.1.2

In the current analysis, we put independent priors on θ1 and θ2. Do you think this is plausible? How would you change the model to take into account the possible dependence? How would this affect the outcome of the Bayesian test?

```{r 9.1.2, echo=FALSE}
cat('# Pledgers
model{
  # Rates and Difference
  theta1 ~ dbeta(1,1)
  theta2 ~ dbeta(1,1)
  delta <- theta1-theta2  
  # Data
  s1 ~ dbin(theta1,n1)
  s2 ~ dbin(theta2,n2)
  # Prior Sampling
  theta1prior ~ dbeta(1,1)
  theta2prior ~ dbeta(1,1)
  deltaprior <- theta1prior-theta2prior
}',file={f<-tempfile()})



data  <- list("s1","s2","n1","n2") # to be passed on to JAGS

myinits <- list(
  list(theta1 = runif(1), theta2 = runif(1), theta1prior = runif(1), theta2prior = runif(1)),
  list(theta1 = runif(1), theta2 = runif(1), theta1prior = runif(1), theta2prior = runif(1)),
  list(theta1 = runif(1), theta2 = runif(1), theta1prior = runif(1), theta2prior = runif(1)))

parameters <- c("theta1", "theta2", "delta", "deltaprior")

# The following command calls JAGS with specific options.
# For a detailed description see the R2jags documentation.
samples <- jags(data, inits=myinits, parameters,
     		model.file =f,
	 			n.chains=3, n.iter=10000, n.burnin=1000, n.thin=1, DIC=T)
# Now the values for delta and deltaprior are in the "samples" object, ready for inspection.

######################################################
# H1: delta is unrestricted
######################################################

# Collect posterior samples across all chains:
df <-data.frame(delta.post=samples$BUGSoutput$sims.list$delta,
                delta.prior=samples$BUGSoutput$sims.list$deltaprior)

#============ BFs based on logspline fit ===========================
fit.prior     <- logspline(df$delta.prior, lbound=-1, ubound=1) # note the bounds.
fit.posterior <- logspline(df$delta.post, lbound=-1, ubound=1)

# 95% confidence interval:
x0 <- qlogspline(0.025,fit.posterior)
x1 <- qlogspline(0.975,fit.posterior)

posterior     <- dlogspline(0, fit.posterior) # this gives the pdf at point delta = 0
prior         <- dlogspline(0, fit.prior)     # based on the logspline fit
BF01          <- posterior/prior

a<-ggplot(data=df)+
  geom_density(aes(x=delta.post),linetype='dashed')+
  geom_density(aes(x=delta.prior))+
  geom_point(x=0,y=prior)+
  geom_point(x=0,y=posterior,color='red')+
  theme_bw()+
  xlim(-1,1)+
  xlab('Delta')

b<-ggplot(data=df)+
  geom_density(aes(x=delta.post),linetype='dashed')+
  geom_density(aes(x=delta.prior))+
  geom_point(x=0,y=prior)+
  geom_point(x=0,y=posterior,color='red')+
  theme_bw()+
  xlim(-0.05,0.05)+
  xlab('Delta')

grid.arrange(a,b,nrow=1)
```

Bayes factor is `r BF01` or 1 / `r BF01`


## 9.2 Order-restricted equality of proportions
Whether pledgers are less likely than non-pledgers to use condoms has a natural interpretation as involving the order-restriction that the rate is lower for the pledgers than for the non-pledgers. 

 This is the same sort of restriction involved in the two-country quiz example in Section 6.4, where it was addressed by using theta2 ∼ dunif(0,1) and theta1 ∼ dunif(0,theta2). While this approximate approach worked satisfactorily for inferring posterior distributions of parameters, an exact approach is needed in the current context of model comparison. This is because the approximate method does not generate a uniform prior distribution of θ1 and θ2 over the region of the joint parameter space satisfying the order-restriction, and model selection, unlike parameter estimation, is likely to be sensitive to this mismatch between available information and the implementation of the model.


```{r 9.2, echo=F}
# Pledgers, Order Constrained Rates
cat('model{
  # Order Constrained Rates
  thetap[1:2] ~ dmnorm(mu[],TI[,])
  theta1 <- phi(cos(angle)*thetap[1]-sin(angle)*abs(thetap[2]))
  theta2 <- phi(sin(angle)*thetap[1]+cos(angle)*abs(thetap[2]))
  
  # Data
  s1 ~ dbin(theta1,n1)
  s2 ~ dbin(theta2,n2)
  
  # Difference
  delta <- theta1-theta2  
  
  # Prior Sampling
  thetapprior[1:2] ~ dmnorm(mu[],TI[,])
  theta1prior <- phi(cos(angle)*thetapprior[1]-sin(angle)*abs(thetapprior[2]))
  theta2prior <- phi(sin(angle)*thetapprior[1]+cos(angle)*abs(thetapprior[2]))
  deltaprior  <- theta1prior-theta2prior
  
  # Constants
  angle <- 45*3.1416/180
  TI[1,1] <- 1
  TI[1,2] <- 0
  TI[2,1] <- 0
  TI[2,2] <- 1
  mu[1] <- 0
  mu[2] <- 0
}',file={f<-tempfile()})


s1 <- 424
s2 <- 5416
n1 <- 777
n2 <- 9072

data  <- list("s1","s2","n1","n2") # to be passed on to JAGS

myinits <- list(
  list(thetap = c(-.8,-.4)),
  list(thetap = c(-.5,-.25)),
  list(thetap = c(-.2,-.1)))

parameters <- c("delta", "deltaprior")

# The following command calls JAGS with specific options.
# For a detailed description see the R2jags documentation.
samples <- jags(data, inits=myinits, parameters,
   			model.file=f,
	 			n.chains=3, n.iter=20000, n.burnin=1000, n.thin=1, DIC=T)

# Collect posterior samples across all chains:
df <-data.frame(delta.post=samples$BUGSoutput$sims.list$delta,
                delta.prior=samples$BUGSoutput$sims.list$deltaprior)

#============ BFs based on logspline fit ===========================
fit.posterior <- logspline(df$delta.post, lbound=-1, ubound=0)

# 95% confidence interval:
x0 <- qlogspline(0.025,fit.posterior)
x1 <- qlogspline(0.975,fit.posterior)

posterior     <- dlogspline(0, fit.posterior) # this gives the pdf at point delta = 0
prior         <- 2*dlogspline(0, fit.prior)     # based on the logspline fit
BF01          <- posterior/prior

a<-ggplot(data=df)+
  geom_density(aes(x=delta.post),linetype='dashed')+
  geom_density(aes(x=delta.prior))+
  geom_point(x=0,y=prior)+
  geom_point(x=0,y=posterior,color='red')+
  theme_bw()+
  xlim(-1,0)+
  xlab('Delta')

b<-ggplot(data=df)+
  geom_density(aes(x=delta.post),linetype='dashed')+
  geom_density(aes(x=delta.prior))+
  geom_point(x=0,y=prior)+
  geom_point(x=0,y=posterior,color='red')+
  theme_bw()+
  xlim(-0.1,0)+
  xlab('Delta')

grid.arrange(a,b,nrow=1)

```

Bayes factor is `r BF01` or 1 / `r BF01`

### Exercise 9.2.1 

Consider an order-restricted test of H0 : δ = 0 versus H3 : δ > 0. What do you think the result will be? Check your intuition by implementing the appropriate graphical model, and estimating the Bayes factor.

{r 9.2.1, echo=F}
# Pledgers, Order Constrained Rates
cat('model{
  # Order Constrained Rates
  thetap[1:2] ~ dmnorm(mu[],TI[,])
  theta1 <- phi(cos(angle)*thetap[1]-sin(angle)*abs(thetap[2]))
  theta2 <- phi(sin(angle)*thetap[1]+cos(angle)*abs(thetap[2]))
  
  # Data
  s1 ~ dbin(theta1,n1)
  s2 ~ dbin(theta2,n2)
  
  # Difference
  delta <- theta1-theta2  
  
  # Prior Sampling
  thetapprior[1:2] ~ dmnorm(mu[],TI[,])
  theta1prior <- phi(cos(angle)*thetapprior[1]-sin(angle)*abs(thetapprior[2]))
  theta2prior <- phi(sin(angle)*thetapprior[1]+cos(angle)*abs(thetapprior[2]))
  deltaprior  <- theta1prior-theta2prior
  
  # Constants
  angle <- 45*3.1416/180
  TI[1,1] <- 1
  TI[1,2] <- 0
  TI[2,1] <- 0
  TI[2,2] <- 1
  mu[1] <- 0
  mu[2] <- 0
}',file={f<-tempfile()})


s1 <- 424
s2 <- 5416
n1 <- 777
n2 <- 9072

data  <- list("s1","s2","n1","n2") # to be passed on to JAGS

myinits <- list(
  list(thetap = c(-.8,-.4)),
  list(thetap = c(-.5,-.25)),
  list(thetap = c(-.2,-.1)))

parameters <- c("delta", "deltaprior")

# The following command calls JAGS with specific options.
# For a detailed description see the R2jags documentation.
samples <- jags(data, inits=myinits, parameters,
     		model.file=f,
	 			n.chains=3, n.iter=20000, n.burnin=1000, n.thin=1, DIC=T)

# Collect posterior samples across all chains:
df <-data.frame(delta.post=samples$BUGSoutput$sims.list$delta,
                delta.prior=samples$BUGSoutput$sims.list$deltaprior)

#============ BFs based on logspline fit ===========================
fit.posterior <- logspline(df$delta.post, lbound=0, ubound=1)

# 95% confidence interval:
x0 <- qlogspline(0.025,fit.posterior)
x1 <- qlogspline(0.975,fit.posterior)

posterior     <- dlogspline(0, fit.posterior) # this gives the pdf at point delta = 0
prior         <- 2*dlogspline(0, fit.prior)     # based on the logspline fit
BF01          <- posterior/prior

a<-ggplot(data=df)+
  geom_density(aes(x=delta.post),linetype='dashed')+
  geom_density(aes(x=delta.prior))+
  geom_point(x=0,y=prior)+
  geom_point(x=0,y=posterior,color='red')+
  theme_bw()+
  xlim(0,1)+
  xlab('Delta')

b<-ggplot(data=df)+
  geom_density(aes(x=delta.post),linetype='dashed')+
  geom_density(aes(x=delta.prior))+
  geom_point(x=0,y=prior)+
  geom_point(x=0,y=posterior,color='red')+
  theme_bw()+
  xlim(0,0.1)+
  xlab('Delta')

grid.arrange(a,b,nrow=1)



